\bfseries How do sailboats work? A simple momentum-transport model

### 2024-08-07 paper/main.tex
Updates from Overleaf

% to-do items:
% ------------
% - Update figure captions to match new figures!
% - Add to the abstract and the introduction and discussion that this is only a 2-D model of sailing. Not 3-D!
% - Add to the abstract the point that the sailboat is defined by 3 dimensionless numbers: The sail-to-keel ratio, and two lift ratios.
% - Change citations to Tao to make it clear that the Tao model is (probably) wrong!!
% - Make sure I cite {symmetry} appropriately in various places. Probably also point out that it is a fun longer read.
% - Make sure we aren't claiming that the model is NOVEL. It is just SIMPLE.
% - Reconsider the post-equation commas? (HWR SAY).
% — Make some comments on the situations in which the ram-pressure force approximation is more appropriate---gale force or lull?
% - Is the latex footnote-footnote separation fucked?
% - Do we still mention "tensors" anywhere? make sure the abstract is consistent with the body.
% - Respond to comments by Hyung Don Ryoo, in email to Hogg, 2021-07-15.
% - Make sure the spirit of the bibliographic comments get pulled into the main text (and removed from the bibiliography).
% - Fix the et al typography in the bibliography?
% - Are we clear that we also find that you can sail faster than the wind, in general, like cross-wind? Emphasize that when we first figure it out.
% - What claims in the literature exist about maximum boat speeds? Have I seen a 2x wind claim? We can beat that. Have I seen a "never faster downwind than the wind" (as in Kimball) or than 0.5 times the wind claim? We can perhaps beat that too.
% - Should we be more careful to use Galilean-relativistic language, like instead of "upwind" say "upwind relative to the water"? Like always?
% - Mention sailing hydrofoil where we discuss the keel lift ratio: The point of the hyrdrofoil is to reduce A_below.
% - Explain why amateur sailboats have huge keels and professional boats have small keels. Upwind sailing is the answer!
% - Kleban points about non-trivial wind force f(theta) and references (in email, maybe).
% - Note that better sailing (tm) is almost never in the direction of the target!
% - Is there an analog of the Betts or Betz limit? Also just generally discussion of power, power dissipated, and so on.
% - Do we adequately mention sustainability in the introduction?
% - How to plot the phase space of dimensionless ratios that allow for UW and DWFTTW?
% - Can we address stability and uniqueness of the solutions to the (quartic, actually) equations?
% - Do we clearly define the dimensionless ratios that define a boat? There are three.
% - Kleban wants polars -- plots that are used in the business. Can do!

@@

We (the authors) are not sailors!
However, we believe that sailing is a beautiful technology.
The boat is powered by the wind, and yet it can travel (possibly with a tacking course) in any direction.
Of course it's not consistent with Galilean relativity\footnote{We refer to Galilean relativity repeatedly in this \documentname.
Of course Galilean relativity is just the low-velocity approximation to special relativity \cite{sr}.
But we are not going to consider air--water velocity differences that are an appreciable fraction of the speed of light (except maybe in the discussion in \secref{sec:discussion}).} to say that the boat is powered by the wind:
There is no absolute velocity, and there is a valid rest frame in which the wind velocity is zero.
The Galilean-relativistic thing to say is that the boat is powered by the relative velocity between the air and the water.
Indeed, the symmetry between the air and the water in the mechanisms of sailing is the fundamental theme of one of the better books on the physics of sailing \cite{symmetry}.
Here we build the simplest possible physical model for a sailboat---and sailing---that is consistent with symmetries and captures all of the leading-order features of the technology.

@@

@@

KLEBAN SAY: Should we quantitatively report on the conditions for dwfttw? And the conditions for upwind? Right now this is all qualitative.
HOGG SAY: YES; THERE are settings for the dimensionless ratios that permit and don't permit it. Not sure how to present this because there are three dimensionless ratios.

@@

Speaking of misconceptions: These results probably do indeed show that you can travel directly downwind faster than the wind.
Of course our sailboat can't (it has to tack), but a pair of sailboats working in tandem could tack oppositely downwind and pull a cargo directly downwind faster than the wind.
That finding, perhaps, resolves some folk conflicts or arguments that it is physically impossible?
In practice it might be hard to build a boat with good enough dimensionless ratios (see \secref{design}).
Maybe it would be easier to build ice boats with the required ratios.
This might be why some sources say it is impossible (KLEBAN CITE).

>>>>

% to-do items:
% ------------
% - Update figure captions to match new figures!
% - Add to the abstract and the introduction and discussion that this is only a 2-D model of sailing. Not 3-D!
% - Add to the abstract the point that the sailboat is defined by 3 dimensionless numbers: The sail-to-keel ratio, and two lift ratios.
% - Change citations to Tao to make it clear that the Tao model is (probably) wrong!!
% - Make sure I cite {symmetry} appropriately in various places. Probably also point out that it is a fun longer read.
% - Make sure we aren't claiming that the model is NOVEL. It is just SIMPLE.
% - Reconsider the post-equation commas? (HWR SAY).
% — Make some comments on the situations in which the ram-pressure force approximation is more appropriate---gale force or lull?
% - Is the latex footnote-footnote separation fucked?
% - Do we still mention "tensors" anywhere? make sure the abstract is consistent with the body.
% - Respond to comments by Hyung Don Ryoo, in email to Hogg, 2021-07-15.
% - Make sure the spirit of the bibliographic comments get pulled into the main text (and removed from the bibiliography).
% - Fix the et al typography in the bibliography?
% - Are we clear that we also find that you can sail faster than the wind, in general, like cross-wind? Emphasize that when we first figure it out.
% - What claims in the literature exist about maximum boat speeds? Have I seen a 2x wind claim? We can beat that. Have I seen a "never faster downwind than the wind" (as in Kimball) or than 0.5 times the wind claim? We can perhaps beat that too.
% - Should we be more careful to use Galilean-relativistic language, like instead of "upwind" say "upwind relative to the water"? Like always?
% - Mention sailing hydrofoil where we discuss the keel lift ratio: The point of the hyrdrofoil is to reduce A_below.
% - Explain why amateur sailboats have huge keels and professional boats have small keels. Upwind sailing is the answer!
% - Kleban points about non-trivial wind force f(theta) and references (in email, maybe).
% - Note that better sailing (tm) is almost never in the direction of the target!
% - Is there an analog of the Betts or Betz limit? Also just generally discussion of power, power dissipated, and so on.
% - Do we adequately mention sustainability in the introduction?
% - How to plot the phase space of dimensionless ratios that allow for UW and DWFTTW?
% - Can we address stability and uniqueness of the solutions to the (quartic, actually) equations?
% - Do we clearly define the dimensionless ratios that define a boat? There are three.
% - Kleban wants polars -- plots that are used in the business. Can do!
% - Discuss "velocity made good" in the text and maybe name-check it in the abstract?

@@

We (the authors) are not serious sailors.
However, we believe that sailing is a beautiful technology.
The boat is powered by the wind, and yet it can travel (possibly with a tacking course) in any direction.
Of course it's not consistent with Galilean relativity\footnote{We refer to Galilean relativity repeatedly in this \documentname.
Of course Galilean relativity is just the low-velocity approximation to special relativity \cite{sr}.
But we are not going to consider air--water velocity differences that are an appreciable fraction of the speed of light (except maybe in the discussion in \secref{sec:discussion}).} to say that the boat is powered by the wind:
There is no absolute velocity, and there is a valid rest frame in which the wind velocity is zero.
The Galilean-relativistic thing to say is that the boat is powered by the relative velocity between the air and the water.
Indeed, the symmetry between the air and the water in the mechanisms of sailing is the fundamental theme of one of the better books on the physics of sailing \cite{symmetry}.
Here we build the simplest possible physical model for a sailboat---and sailing---that is consistent with symmetries and captures all of the leading-order features of the technology.

@@

HOGG: Argument about the wedge being squeezed.

HOGG: Diagrams.

@@

KLEBAN SAY: Should we quantitatively report on the conditions for dwfttw? And the conditions for upwind? Right now this is all qualitative.
HOGG SAY: YES; THERE are settings for the dimensionless ratios that permit and don't permit it. Not sure how to present this because there are three dimensionless ratios. Working on it.

@@

Speaking of misconceptions: These results probably do indeed show that you can travel directly downwind faster than the wind.
Of course our sailboat can't (it has to tack), but a pair of sailboats working in tandem could tack oppositely downwind and pull a cargo directly downwind faster than the wind.
That finding, perhaps, resolves some folk conflicts or arguments that it is physically impossible?
In practice it might be hard to build a boat with good enough dimensionless ratios (see \secref{sec:design}).
Maybe it would be easier to build ice boats with the required ratios.
This might be why some sources say it is impossible (KLEBAN CITE).
